Finiteness in Cubical Type Theory

### 2020-01-21 paper.tex
cleaned etc

  It is defined as a Higher Inductive Type:
  \begin{equation} {
    \begin{alignat*}{3}
      \lVert A \rVert \coloneqq & \; \lvert \wc \rvert &&: A \rightarrow \lVert A \rVert ; \\
                              | & \; \text{squash}     &&: \Pi {(x, y : \lVert A \rVert)} , x \equiv y  ; 
    \end{alignat*} }
  \end{equation}
  We will use three eliminators from \(\lVert A \rVert\) in this paper.
  \begin{enumerate}
  \item \label{elim-prop-prop} For any function \(A \rightarrow B\), where
    \(\text{isProp}(B)\), we have a function \(\lVert A \rVert \rightarrow B\).
  \item \label{elim-prop-monad} A special case of \ref{elim-prop-prop} implies
    that \(\lVert \cdot \rVert\) forms a monad: this means that we get the usual
    Monadic operators on \(\lVert \cdot \rVert\) (bind, pure, fmap, etc.).
    \item \label{elim-prop-coh} We can eliminate from \(\lVert A \rVert\) with a
      function \(f : A \rightarrow B\) iff \(f\) ``doesn't care'' about the
      choice of \(A\) (\(\Pi {(x , y : A)} , f(x) \equiv f(y) \)).
      Formally speaking, \(f\) needs to be ``coherently constant''
      \cite{krausGeneralUniversalProperty2015}, and \(B\) needs to be an
      \(n\)-type for some finite \(n\).
  \end{enumerate}
\end{romdefinition}
\section{Finiteness Predicates}
In this section, we will define and briefly describe each of the five predicates
in Figure~\ref{finite-classification}.
The reason we explore predicates other than our focus (cardinal finiteness) is
that we can often prove things like closure much more readily on the simpler
predicates.
The relations (which we will prove in the next section) then allow us to
transfer those proofs onto Kuratowski finiteness.
\subsection{Split Enumerability}
\begin{romdefinition}[Split Enumerable Set] \label{split-enum-def}
  \begin{equation} \label{split-enum-def-eqn}
    \mathcal{E}!(A) \coloneqq \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \in \mathit{xs}
  \end{equation}
  We call the first component of this pair the ``support'' list, and the second
  component the ``cover'' proof.
  An equivalent version of this predicate was called \verb+Listable+ in
  \cite{firsovDependentlyTypedProgramming2015}.
\end{romdefinition}

We tend to prefer list-based definitions of finiteness, rather than ones based
on bijections or surjections.
This is purely a matter of perspective, however: the definition above is
precisely equivalent to a split surjection from a finite prefix of the natural
numbers.
\begin{romlemma} \label{split-enum-is-split-surj}
  \begin{equation}
    \mathcal{E}!(A) \simeq \Sigma (n : \mathbb{N}) , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
  \end{equation}
\end{romlemma}
\begin{proof}
  \begin{align*}
    \mathcal{E}!(A)
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \in \mathit{xs}
    && \text{Def.~\ref{split-enum-def} }(\mathcal{E}!)
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{fib}_{\text{snd}(\mathit{xs})}(x)
    && \text{Eqn.~\ref{container-membership} } (\in)
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{sp-surj}(\text{snd}(\mathit{xs}))
    && \text{Eqn.~\ref{sp-surj-eqn} (sp-surj)}
    \\
    & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{sp-surj}(\text{snd}(\mathit{xs}))
    && \text{Def.~\ref{List} } (\mathbf{List})
    \\
    & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{sp-surj}(\text{snd}(\mathit{xs}))
    && \text{Eqn.~\ref{container-interp} } (\llbracket \wc \rrbracket)
    \\
    & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{sp-surj}(f)
    && \text{Reassociation of } \Sigma
    \\
    & \simeq \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \twoheadrightarrow ! \; A )
    && \text{Eqn.~\ref{sp-surj-arrow-eqn} } (\twoheadrightarrow!)
  \end{align*}
  \qed
\end{proof}

@@

Where split enumerability was the enumeration form of a split surjection from
\(\mathbf{Fin}\), manifest Bishop finiteness is the enumeration form of an
\emph{equivalence} with \(\mathbf{Fin}\).
\begin{romlemma} \label{bishop-equiv}
  \begin{equation}
    \mathcal{B}(A) \simeq \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \simeq A \right)
  \end{equation}
\end{romlemma}
\begin{proof}
  \begin{align*}
    \mathcal{B}(A)
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \inunique \mathit{xs}
    && \text{Def.~\ref{bish-def} } (\mathcal{B})
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{isContr} \left(\text{fib}_{\text{snd}(\mathit{xs})}(x)\right)
    && \text{Def.~\ref{uniq-memb-def} } (\inunique)
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    && \text{Eqn.~\ref{is-equiv-def} (isEquiv)}
    \\
    & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    && \text{Def.~\ref{List} } (\mathbf{List})
    \\
    & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    && \text{Eqn.~\ref{container-interp} } (\llbracket \wc \rrbracket)
    \\
    & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{isEquiv}(f)
    && \text{Reassociation of } \Sigma
    \\
    & \simeq \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \simeq \; A )
    && \text{Eqn.~\ref{surj-arrow-eqn} } (\twoheadrightarrow)
  \end{align*}
  \qed
\end{proof}
\subsection{Cardinal}
\begin{itemize}
\item Definition
\item Cardinality
\end{itemize}
\subsection{Manifest Enumerable}
\begin{romdefinition}[Manifest Enumerability]
  \begin{equation}
    \mathcal{E}(A) = \Sigma {(\mathit{xs} : \mathbf{List}(A))} , \Pi {(x : A)} , \lVert x \in \mathit{xs} \rVert
  \end{equation}
\end{romdefinition}
Again, the only difference with this type and split enumerability is the
membership proof: here we have propositionally truncated it.
This has two effects.
First, it means that this proof represents a true surjection (rather than a
split surjection) from \(\mathbf{Fin}\).
\begin{romlemma}
  \begin{equation}
    \mathcal{E}(A) \simeq \Sigma(n : \mathbb{N}) , (\mathbf{Fin}(n) \twoheadrightarrow A)
  \end{equation}
\end{romlemma}
\begin{proof}
  \todo[inline]{Proof}
\end{proof}

Secondly, it means the predicate does not imply decidable equality.
More significantly, it allows the predicate to be defined over non-set types,
like the circle.
\begin{romlemma}
  The circle \(S^1\) is manifestly enumerable.
\end{romlemma}
\begin{proof}
  \todo[inline]{Proof}
\end{proof}

\subsection{Kuratowski Finiteness}
Much work has already been done on Kuratowski finiteness in HoTT in
\cite{fruminFiniteSetsHomotopy2018}.
As a result, we will not needlessly repeat proofs, rather we will just give a
brief introduction to the topic and point out where our treatment differs.

The first thing we must define is a representation of subsets.
\begin{romdefinition}[Kuratowski Finite Subset]
  \(\mathcal{K}(A)\) is the type of Kuratowski-finite subsets of \(A\).
  \begin{equation}
    \begin{aligned}
      \mathcal{K}(A) =&
      \; [] : \mathcal{K}(A) ; \\
      \vert& \; \wc \dblcolon \wc : A \rightarrow \mathcal{K}(A) \rightarrow \mathcal{K}(A) ; \\
      \vert& \; \text{com} : \Pi (x, y: A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon y \dblcolon \mathit{xs} \simeq y \dblcolon x \dblcolon \mathit{xs} ; \\
      \vert& \; \text{dup} : \Pi (x : A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon x \dblcolon \mathit{xs} \simeq x \dblcolon \mathit{xs} ; \\
      \vert& \; \text{trunc} : \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \simeq \mathit{ys}) , p \simeq q ;
    \end{aligned}
  \end{equation}
\end{romdefinition}
The Kuratowski finite subset is a free join semilattice (or, equivalently, a
free commutative idempotent monoid).
More prosaically, \(\mathcal{K}\) is the abstract data type for finite sets, as
defined in the Boom hierarchy \cite{boomFurtherThoughtsAbstracto1981,
  bunkenburgBoomHierarchy1994}.
However, rather than just being a specification, \(\mathcal{K}\) is fully usable
as a data type in its own right, thanks to HITs.

@@

Next, we need a way to say that an entire type is Kuratowski finite.
For that, we will need to define membership of \(\mathcal{K}\).
\begin{romdefinition}[Membership of \(\mathcal{K}\)]
  Membership is defined by pattern-matching on \(\mathcal{K}\).
  The two point constructors are handled like so:
  \begin{equation}
    \begin{alignedat}{2}
      x \in&& \; []                      &= \bot \\
      x \in&& \; y \dblcolon \mathit{ys} &= \lVert x \simeq y \uplus x \in \mathit{ys} \rVert
    \end{alignedat}
  \end{equation}
  The \(\text{com}\) and \(\text{dup}\) constructors are handled by proving that
  the truncated form of \(\uplus\) is itself commutative and idempotent.
  The type of propositions is itself a set, satisfying the \(\text{trunc}\)
  constructor.
\end{romdefinition}
Finally, we have enough background to define Kuratowski finiteness.
\begin{romdefinition}[Kuratowski Finiteness]
  \begin{equation}
    \mathcal{K}^{f}(A) = \Sigma {(\mathit{xs} : \mathcal{K}(A))} , \Pi (x : A) , x \in \mathit{xs}
  \end{equation}
\end{romdefinition}

We also have the following two lemmas, proven in both
\cite{fruminFiniteSetsHomotopy2018} and our formalisation.
\begin{romlemma}
  \(\mathcal{K}^f\) is a mere proposition.
\end{romlemma}
\begin{romlemma}
  This circle \(S^1\) is Kuratowski finite.
\end{romlemma}
The second of these in particular tells us that the ``Kuratowski-finite types''
are not necessarily sets; it also tells us that we cannot derive decidable
equality from a proof of Kuratowski finiteness.
\section{Relations Between Each Finiteness Definition}
\begin{itemize}
  \item SE -> MB (and back)
  \item MB <-> ME
  \item MB <-> C
  \item C <-> K
  \item ME -> K
\end{itemize}
\section{Topos}
\begin{itemize}
  \item \(\bot\)
  \item \(\top\)
  \item Bool
  \item \(\Sigma\)
  \item \(\Pi\)
  \item Pre-categories to categories
  \item \(\Pi\)-pretopos
\end{itemize}
\section{Countably Infinite Types}
\begin{itemize}
  \item Definition
  \item \(\Sigma\)
  \item Kleene star
  \item Truncated
  \item Topos
\end{itemize}
\section{Search}
\begin{itemize}
  \item Omniscience
  \item Search
  \item Curry / uncurry
  \item Synth function
\end{itemize}

% % Taking finiteness as an example again, if we prove that some type is finite via
% % bijection, in MLTT such a proof would have to reveal the particular bijection.
% % This becomes problematic when defining, for instance, the category of finite
% % sets: the objects of this category would ideally be types coupled with proofs of
% % their finiteness, but in order to

% % The five in particular that we are interested in are organised in
% % figure~\ref{finite-classification}.
% % We especially want to examine what separates each predicate: what extra
% % information do you need, for instance, to go from a proof of Cardinal finiteness
% % (section~\ref{cardinal}) to manifest Bishop finiteness
% % (section~\ref{manifest-bishop}).

% % Next, we will examine the category-theoretic and topos-theoretic interpretations
% % of each of the predicates.
% % We will give a ``category of finite sets'' for each finiteness predicate, and
% % then prove the various closure properties required to arrive at the relevant
% % topos (we show that decidable Kuratowski finite sets form a \(\Pi\)-pretopos,
% % for instance).

% % ways to prove finiteness end up proving the same
% % thing: constructively, however, each defines a fundamentally different
% % ``version'' of finiteness.
% % In Martin-Löf type theory \cite{martin-lofIntuitionisticTypeTheory1980}, these
% % variations (and more) have been explored in
% % \cite{spiwackConstructivelyFinite2010} and
% % \cite{firsovVariationsNoetherianness2016}.

% % Homotopy Type Theory \cite{hottbook} further adds to this story in two ways.
% % First, its universe of types includes non-sets: whether or not a finiteness
% % predicate includes these gives us another dimension on which to discriminate.

% % The second addition is concerned with the hiding of information.
% % Occasionally in a constructive setting you don't want to have to reveal how you
% % proved something.
% % To model logical disjunction, for instance, we usually use a disjoint union: the
% % model isn't entirely correct, though, because a proof of \(A \text{or} B\)
% % reveals which of \(A\) or \(B\) we actually proved.

% % logical disjunction, for instance, is usually modelled with a
% % disjoint union.
% % This always means, however, that to prove ``A or B'' you have to reveal which of
% % A or B you did prove.

% % The other headline feature of HoTT's generalised equality is univalence: put
% % simply, it allows us to treat isomorphic types as equal, giving us in the
% % constructive world access to a technique that is commonplace classically.
% % It turns out that finite types have a lot to say about isomorphisms, and we will
% % rely heavily on this aspect of HoTT to work with them.
% % \subsection{Contributions}
% % In this work, we will explore finite types in Cubical Type Theory
% % \cite{cohenCubicalTypeTheory2016}, and expose their relationship to infinite
% % types, species, and demonstrate their practical uses for proofs in dependently
% % typed programming languages.

% % In section~\ref{strong-finiteness}, we will explore the ``strong'' notions of
% % finiteness (i.e. those at least as strong as Kuratowski finiteness
% % \cite{kuratowskiNotionEnsembleFini1920}), with a special focus on cardinal
% % finiteness.
% % We will delineate some subtly different finiteness predicates, and prove the
% % relationship between each.
% % We will prove, then, that every discrete Kuratowski finite type is cardinal
% % finite, and subsequently will show that discrete Kuratowski finite sets form a
% % topos.

% % In section~\ref{species-and-containers}, we will redefine our finiteness
% % predicates using \emph{containers}, and prove that the new definitions are
% % equivalent to the old.
% % Using this, we will prove much stronger relationships between our finiteness
% % predicates and relation-based definitions of finiteness.
% % Finally, We will show how the species-container duality falls out naturally from
% % these new definitions.

% % In section~\ref{infinite-cardinalities}, we will extend our study of finite
% % types to infinite but countable types.
% % We will see that the finiteness predicates are mirrored with countable
% % counterparts, and we will prove closure under the Kleene star and plus.

% % Proofs of finiteness have well-known practical applications in
% % constructive mathematics \cite{firsovDependentlyTypedProgramming2015}.
% % In section~\ref{practical}, we build a library which exploits these uses in
% % Cubical Agda \cite{vezzosiCubicalAgdaDependently2019}, allowing automation of
% % complex proofs over finite types.
% % We frame this in terms of the principle of omniscience for finite types.
% % Thanks to the flexibility afforded to us by Cubical Type Theory, we are able to
% % go further than the usual examples of this kind of proof automation: as well as
% % proving properties about functions, we can synthesise functions whole-cloth from
% % their desired properties.
% % Through the unified interface for finite and countable types, we can
% % reuse the automation machinery for \emph{partial} proof search over infinite
% % search spaces.
% % Along the way, we extend the work in \cite{allaisGenericLevelPolymorphic2019} to
% % prove isomorphisms between the curried and uncurried forms of \(n\)-ary
% % dependent functions.

% % Classically, of course, each of these statements are equivalent.
% % Constructively, though, all four predicates are fundamentally different.

% % Part of the difference comes from ``leaking'' information about the finite type
% % in the proof of its finiteness.
% % If I demonstrate a type is finite by showing a surjection to that type, I've
% % revealed a way to construct all of the elements of my type.
% % If I show that it is bounded, I reveal an upper bound on the cardinality of the
% % type.

% \subsection{Strong Finiteness Predicates}
% We will first explore the ``strong'' notions of finiteness (i.e. those at least
% as strong as Kuratowski finiteness \cite{kuratowskiNotionEnsembleFini1920}),
% with a special focus on cardinal finiteness (section~\ref{cardinal}), and
% manifest enumerability (section~\ref{manifest-enumerability}), which is new, to
% our knowledge.

% Figure~\ref{finite-classification} organises the predicates according to
% their strength.
% We will go through each of the predicates, proving how to weaken each (i.e. we
% will provide a proof that every cardinally finite type is Kuratowski finite),
% and how to strengthen them, given the required property.
% In terms of figure~\ref{finite-classification}, this amounts to providing proofs
% for each arrow.
% Our main proof in this regard is that we can derive manifest Bishop finiteness
% from Kuratowski finiteness plus a total order.

% We will---through the use of containers
% \cite{abbottContainersConstructingStrictly2005}---formally prove the equivalence
% these predicates have with the usual function relations i.e. we will show that a
% proof of manifest enumerability is precisely equivalent to a surjection from a
% finite prefix of the natural numbers.

% For each predicate, we will also prove its closure over sums and products in
% both dependent and non-dependent forms, if such a closure exists.
% This will culminate in our main result for this section: the formal proof that
% decidable Kuratowski finite sets form a topos.
% \subsection{Infinite Types}
% In section~\ref{infinite-cardinalities}, we will extend our study of finite
% types to infinite but countable types.
% We will see that the classification of finiteness predicates mirrors that of
% their countable counterparts, and we will prove closure under the Kleene star.

% \subsection{Practical Uses}
% Finally, in section~\ref{practical} we will demonstrate some practical uses of
% the finiteness proofs in Cubical Agda.
% We will show how to use well-known techniques
% \cite{firsovDependentlyTypedProgramming2015, devrieseBrightSideType2011} to
% automate proofs of functions with finite (and infinite) domains.
% We further show how to automate the synthesis of functions from desired
% properties.
% \section{Split Enumerability} \label{split-enumerable}
% We will start with the simplest definition of finiteness: we say a set is
% enumerable if there is a list of its elements which contains every element in
% the set.

% Before giving the definition of the predicate, we will first define lists (and
% membership thereof): we have chosen a \emph{container}-based definition for this
% work.

% Internally, in our formalisation, we actually use the standard
% inductive definition of lists more often (it tends to work better in more
% complex algorithms, and functions on it seems to satisfy the termination
% checker more readily).
% However, since both types are equivalent, univalence allows us to transport to
% whichever representation is more convenient in a given situation.
% For the higher-level proofs we present here, though, the container-based
% definition greatly simplifies certain steps, which is why we have chosen it as
% our representation.

>>>>

  It is defined as a Higher Inductive Type:
  \begin{equation} {
    \begin{alignat*}{3}
      \lVert A \rVert \coloneqq & \; \lvert \wc \rvert &&: A \rightarrow \lVert A \rVert ; \\
                              | & \; \text{squash}     &&: \Pi {(x, y : \lVert A \rVert)} , x \equiv y  ; 
    \end{alignat*} }
  \end{equation}
  We will use three eliminators from \(\lVert A \rVert\) in this paper.
  \begin{enumerate}
  \item \label{elim-prop-prop} For any function \(A \rightarrow B\), where
    \(\text{isProp}(B)\), we have a function \(\lVert A \rVert \rightarrow B\).
  \item \label{elim-prop-monad} A special case of \ref{elim-prop-prop} implies
    that \(\lVert \cdot \rVert\) forms a monad: this means that we get the usual
    Monadic operators on \(\lVert \cdot \rVert\) (bind, pure, fmap, etc.).
    \item \label{elim-prop-coh} We can eliminate from \(\lVert A \rVert\) with a
      function \(f : A \rightarrow B\) iff \(f\) ``doesn't care'' about the
      choice of \(A\) (\(\Pi {(x , y : A)} , f(x) \equiv f(y) \)).
      Formally speaking, \(f\) needs to be ``coherently constant''
      \cite{krausGeneralUniversalProperty2015}, and \(B\) needs to be an
      \(n\)-type for some finite \(n\).
  \end{enumerate}
\end{romdefinition}
\begin{romdefinition}[\(S^1\)] \label{circle-def}
  The circle, \(S^1\), can be represented in HoTT as a higher inductive type.
  \begin{equation}
    \begin{alignat}{5}
      S^1 \coloneqq & \; \text{base} &&: S^1 ; \\
                  | & \; \text{loop} &&: \text{base} \equiv \text{base} ; 
    \end{alignat}
  \end{equation}
  We will use it here as an example of a non-set type, i.e. a type for which not
  all paths are equal.
  This also means that it does not have decidable equality.
\end{romdefinition}
\section{Finiteness Predicates}
In this section, we will define and briefly describe each of the five predicates
in Figure~\ref{finite-classification}.
The reason we explore predicates other than our focus (cardinal finiteness) is
that we can often prove things like closure much more readily on the simpler
predicates.
The relations (which we will prove in the next section) then allow us to
transfer those proofs onto Kuratowski finiteness.
\subsection{Split Enumerability}
\begin{romdefinition}[Split Enumerable Set] \label{split-enum-def}
  \begin{equation} \label{split-enum-def-eqn}
    \mathcal{E}!(A) \coloneqq \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \in \mathit{xs}
  \end{equation}
  We call the first component of this pair the ``support'' list, and the second
  component the ``cover'' proof.
  An equivalent version of this predicate was called \verb+Listable+ in
  \cite{firsovDependentlyTypedProgramming2015}.
\end{romdefinition}

We tend to prefer list-based definitions of finiteness, rather than ones based
on bijections or surjections.
This is purely a matter of perspective, however: the definition above is
precisely equivalent to a split surjection from a finite prefix of the natural
numbers.
\begin{romlemma} \label{split-enum-is-split-surj}
  \begin{equation}
    \mathcal{E}!(A) \simeq \Sigma (n : \mathbb{N}) , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
  \end{equation}
\end{romlemma}
\begin{proof}
  \begin{align*}
    \mathcal{E}!(A)
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \in \mathit{xs}
    && \text{def.~\ref{split-enum-def} }(\mathcal{E}!)
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{fib}_{\text{snd}(\mathit{xs})}(x)
    && \text{eqn.~\ref{container-membership} } (\in)
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{sp-surj}(\text{snd}(\mathit{xs}))
    && \text{eqn.~\ref{sp-surj-eqn} (sp-surj)}
    \\
    & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{sp-surj}(\text{snd}(\mathit{xs}))
    && \text{def.~\ref{List} } (\mathbf{List})
    \\
    & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{sp-surj}(\text{snd}(\mathit{xs}))
    && \text{eqn.~\ref{container-interp} } (\llbracket \wc \rrbracket)
    \\
    & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{sp-surj}(f)
    && \text{Reassociation of } \Sigma
    \\
    & \simeq \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \twoheadrightarrow ! \; A )
    && \text{eqn.~\ref{sp-surj-arrow-eqn} } (\twoheadrightarrow!)
  \end{align*}
  \qed
\end{proof}

@@

Where split enumerability was the enumeration form of a split surjection from
\(\mathbf{Fin}\), manifest Bishop finiteness is the enumeration form of an
\emph{equivalence} with \(\mathbf{Fin}\).
\begin{romlemma} \label{bishop-equiv}
  \begin{equation}
    \mathcal{B}(A) \simeq \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \simeq A \right)
  \end{equation}
\end{romlemma}
\begin{proof}
  \begin{align*}
    \mathcal{B}(A)
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \inunique \mathit{xs}
    && \text{def.~\ref{bish-def} } (\mathcal{B})
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{isContr} \left(\text{fib}_{\text{snd}(\mathit{xs})}(x)\right)
    && \text{def.~\ref{uniq-memb-def} } (\inunique)
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    && \text{eqn.~\ref{is-equiv-def} (isEquiv)}
    \\
    & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    && \text{def.~\ref{List} } (\mathbf{List})
    \\
    & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    && \text{eqn.~\ref{container-interp} } (\llbracket \wc \rrbracket)
    \\
    & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{isEquiv}(f)
    && \text{Reassociation of } \Sigma
    \\
    & \simeq \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \simeq \; A )
    && \text{eqn.~\ref{surj-arrow-eqn} } (\twoheadrightarrow)
  \end{align*}
  \qed
\end{proof}
\subsection{Cardinal}
Each finiteness predicate so far has contained an \emph{ordering} of the
underlying type.
For our purposes, this is too much information: it means that when constructing
the ``category of finite sets'' later on, instead of each type having one
canonical representative, it will have \(n!\), where \(n\) is the cardinality of
the type.

To remedy the problem, we will use propositional truncation
(def.~\ref{prop-trunc}).
\begin{romdefinition}[Cardinal Finiteness]
  \begin{equation}
    \mathcal{C}(A) \coloneqq \lVert \mathcal{B}(A) \rVert
  \end{equation}
\end{romdefinition}
At first glance, it might seem that we lose any useful properties we could
derive from \(\mathcal{B}\).
Luckily, this is not the case: by eliminator \ref{elim-prop-coh} of
def.~\ref{prop-trunc}, we need only show that the output is uniquely determined.

The following two lemmas are proven in
\cite{yorgeyCombinatorialSpeciesLabelled2014} (Proposition 2.4.9 and 2.4.10,
respectively), in much the same way as we have done here.
Our contribution for this section is simply the formalisation.
\begin{romlemma}
  Given a cardinally finite type, we can derive the type's cardinality, as well
  as a propositionally truncated proof of equivalence with \(\textbf{Fin}\)s of
  the same cardinality.
  \begin{equation}
    \mathcal{C}(A) \rightarrow \Sigma {(n : \mathbb{N})} , \lVert \textbf{Fin}(n) \simeq A \rVert
  \end{equation}
\end{romlemma}
\begin{romlemma} \label{cardinal-finite-discrete}
  Any cardinal-finite set has decidable equality.
\end{romlemma}
\subsection{Manifest Enumerability}
\begin{romdefinition}[Manifest Enumerability]
  \begin{equation}
    \mathcal{E}(A) \coloneqq \Sigma {(\mathit{xs} : \mathbf{List}(A))} , \Pi {(x : A)} , \lVert x \in \mathit{xs} \rVert
  \end{equation}
\end{romdefinition}
As with manifest Bishop finiteness, the only difference with this type and split
enumerability is the membership proof: here we have propositionally truncated
it.
This has two effects.
First, it means that this proof represents a true surjection (rather than a
split surjection) from \(\mathbf{Fin}\).
\begin{romlemma}
  \begin{equation}
    \mathcal{E}(A) \simeq \Sigma(n : \mathbb{N}) , (\mathbf{Fin}(n) \twoheadrightarrow A)
  \end{equation}
\end{romlemma}

Secondly, it means the predicate does not imply decidable equality.
More significantly, it allows the predicate to be defined over non-set types,
like the circle (def.~\ref{circle-def}).
\begin{romlemma}
  The circle \(S^1\) is manifestly enumerable.
\end{romlemma}
\subsection{Kuratowski Finiteness}
Much work has already been done on Kuratowski finiteness in HoTT in
\cite{fruminFiniteSetsHomotopy2018}.
As a result, we will not needlessly repeat proofs, rather we will just give a
brief introduction to the topic and point out where our treatment differs.

The first thing we must define is a representation of subsets.
\begin{romdefinition}[Kuratowski Finite Subset]
  \(\mathcal{K}(A)\) is the type of Kuratowski-finite subsets of \(A\).
  \begin{equation}
    \begin{alignat}{2}
      \mathcal{K}(A) \coloneqq&
             \; []                &&: \mathcal{K}(A) ; \\
      \vert& \; \wc \dblcolon \wc &&: A \rightarrow \mathcal{K}(A) \rightarrow \mathcal{K}(A) ; \\
      \vert& \; \text{com}        &&: \Pi (x, y: A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon y \dblcolon \mathit{xs} \equiv y \dblcolon x \dblcolon \mathit{xs} ; \\
      \vert& \; \text{dup}        &&: \Pi (x : A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon x \dblcolon \mathit{xs} \equiv x \dblcolon \mathit{xs} ; \\
      \vert& \; \text{trunc}      &&: \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \equiv \mathit{ys}) , p \equiv q ;
    \end{alignat}
  \end{equation}

  The first two constructors are point constructors, giving ways to create
  values of type \(\mathcal{K}(A)\).
  They are also recognisable as the two constructors for finite lists, a type
  which represents the free monoid.

  The next two constructors add extra paths to the type: equations that usage of
  the type must obey.
  These extra paths turn the free monoid into the free \emph{commutative} (com)
  \emph{idempotent} (dup) monoid.

  The final constructor enforces that the type \(\mathcal{K}(A)\) must be a set.
\end{romdefinition}
The Kuratowski finite subset is a free join semilattice (or, equivalently, a
free commutative idempotent monoid).
More prosaically, \(\mathcal{K}\) is the abstract data type for finite sets, as
defined in the Boom hierarchy \cite{boomFurtherThoughtsAbstracto1981,
  bunkenburgBoomHierarchy1994}.
However, rather than just being a specification, \(\mathcal{K}\) is fully usable
as a data type in its own right, thanks to HITs.

@@

Next, we need a way to say that an entire type is Kuratowski finite.
For that, we will need to define membership of \(\mathcal{K}\).
\begin{romdefinition}[Membership of \(\mathcal{K}\)]
  Membership is defined by pattern-matching on \(\mathcal{K}\).
  The two point constructors are handled like so:
  \begin{equation}
    \begin{alignedat}{2}
      x \in&& \; []                      &\coloneqq \bot \\
      x \in&& \; y \dblcolon \mathit{ys} &\coloneqq \lVert x \equiv y \uplus x \in \mathit{ys} \rVert
    \end{alignedat}
  \end{equation}
  The \(\text{com}\) and \(\text{dup}\) constructors are handled by proving that
  the truncated form of \(\uplus\) is itself commutative and idempotent.
  The type of propositions is itself a set, satisfying the \(\text{trunc}\)
  constructor.
\end{romdefinition}
Finally, we have enough background to define Kuratowski finiteness.
\begin{romdefinition}[Kuratowski Finiteness]
  \begin{equation}
    \mathcal{K}^{f}(A) = \Sigma {(\mathit{xs} : \mathcal{K}(A))} , \Pi (x : A) , x \in \mathit{xs}
  \end{equation}
\end{romdefinition}

We also have the following two lemmas, proven in both
\cite{fruminFiniteSetsHomotopy2018} and our formalisation.
\begin{romlemma}
  \(\mathcal{K}^f\) is a mere proposition.
\end{romlemma}
\begin{romlemma}
  This circle \(S^1\) is Kuratowski finite.
\end{romlemma}
The second of these in particular tells us that the ``Kuratowski-finite types''
are not necessarily sets; it also tells us that we cannot derive decidable
equality from a proof of Kuratowski finiteness.
\section{Relations Between Each Finiteness Definition}
We now look at under what conditions a given finiteness predicate satisfies
another.
\subsection{Split Enumerability and Manifest Bishop Finiteness}
\subsection{Manifest Bishop Finiteness and Manifest Enumerability}
\subsection{Manifest Bishop Finiteness and Cardinal Finiteness}
\subsection{Manifest Enumerability and Kuratowski Finiteness}
\subsection{Cardinal Finiteness and Kuratowski Finiteness}
\section{Topos}
\begin{itemize}
  \item \(\bot\)
  \item \(\top\)
  \item Bool
  \item \(\Sigma\)
  \item \(\Pi\)
  \item Pre-categories to categories
  \item \(\Pi\)-pretopos
\end{itemize}
\section{Countably Infinite Types}
\begin{itemize}
  \item Definition
  \item \(\Sigma\)
  \item Kleene star
  \item Truncated
  \item Topos
\end{itemize}
\section{Search}
\begin{itemize}
  \item Omniscience
  \item Search
  \item Curry / uncurry
  \item Synth function
\end{itemize}
